Language Use in a Multilingual Tweet Corpus

### 2018-07-19 paper.tex
Language

Our base assumption is that geo-located tweets are a representative and unbiased sample of tweets from Latvia. Thus, an objective collection should exhibit similar properties as the geo-located one. The main property we study in this work is the language proportion. We hypothesize, that an objectively extended collection should have similar language distribution as the initial.

@@

The analysis shows that---despite our attempt---the extended collection is biased toward content in Latvian: proportionally there are more tweets in Latvian in the extended collection than in geo-located. We observe this phenomena not only at the ``global level'' (the whole collection), but also at the ``local level'' (across the topics). \hl{In all topics, Latvian is more dominant than expected.}

@@

Over the period from 15 April 2017 to \hl{16 July 2018} the initial set of \hl{1\,959\,695} tweets was collected from the \texttt{POST status/filter} endpoint of the Twitter Streaming API. Following \cite{milajevs:2017:BUCC}, the \texttt{locations} parameter was set to the bounding box of Riga, the capital of Latvia. It is small enough to fit into Twitter API restrictions and covers about 40\% of the popilation of Latvia.\footnotemark{}
%
\footnotetext{The coordinates of the bounding box are \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}.}
% IMS.2 according to http://worldpopulationreview.com/countries/latvia-population/, the population of Riga is less than half the total population of Latvia.  However, Twitter only allows bounding boxes for location to be so big, so you can't bound the whole country.

@@

To comply with Twitter's terms of service, on \hl{16 July 2018}, the raw tweet data was redownloaded to get rid of deleted tweets. The tweets that originated a retweet were added to the collection. Also, we have noticed a large number of tweets that came from Sweden (probably because of the imprecise \texttt{locations} parameter value), so the tweets which location country code was \texttt{SE} were omitted. This resulted in \hl{1\,415\,984} tweets that formed the final collection presented here.\footnotemark{}

@@

The topics are:\footnotemark{}%
\footnotetext{The topic file is available as \sfile{topics.json.txt}. Relevance judgments are available as \sfile{relevance\_judgments.csv}.}%
\begin{itemize}
\item \texttt{LV001:} \textbf{Latvia 100} \ The Centennial Celebration of the Republic of Latvia.
\item \texttt{LV002:} \textbf{Ice Hockey} \ Ice hockey.
\item \texttt{LV003:} \textbf{Refugees} \ Refugee crisis in Europe, attitude to immigration and immigrants.
\item \texttt{LV004:} \textbf{Brexit} \ Withdrawal of the United Kingdom from the European Union known as Brexit.
\item \texttt{LV005:} \textbf{Olympics} \ The Olympic games and the Latvian team.
\item \texttt{LV006:} \textbf{Winter} \ Winter, snow and cold weather.
\item \texttt{LV007:} \textbf{May 4} \ The Restoration of Independence Day.
\item \texttt{LV008:} \textbf{May 8-9} \ Remembrance of the end of Word War II (May 8). It is also honored as the Victory Day on May 9. The Europe Day is observed on May 9. Tweets about any of the three events are relevant.
\item \texttt{LV009:} \textbf{Midsummer} \ Midsummer celebration on June 23-24.
\item \texttt{LV010:} \textbf{November 11} \ A memorial day for soldiers who fought for the independence of Latvia.
\item \texttt{LV011:} \textbf{November 18} \ Proclamation Day of the Republic of Latvia.
\item \texttt{LV012:} \textbf{Christmas} \ Christmas.
\item \texttt{LV013:} \textbf{New Year} \ New Year.
\item \texttt{LV014:} \textbf{March 16} \ Remembrance Day of the Latvian legionnaires.\footnotemark{}
\item \texttt{LV015:} \textbf{The Chronicles of Melanie} \ Mel\=anijas Hronika (The Chronicles of Melanie) is a Latvian movie that was selected for the Foreign-Language Category for Oscars.
\item \texttt{LV006:} \textbf{Blizzard of Souls} \ A Latvian movie Dv\=ese\c{l}u Putenis (Blizzard of Souls) that is in production by the time of writing.
\end{itemize}

>>>>

Our base assumption is that geo-located tweets are a representative and unbiased sample of tweets from Latvia. Thus, an objective collection should exhibit similar properties to the geo-located one. The main property we study in this work is the language proportion. We hypothesize, that an objectively extended collection should have similar language distribution as the initial collection.

@@

The analysis shows that---despite our attempt---the extended collection is biased toward content in Latvian: proportionally there are more tweets in Latvian in the extended collection than in the geo-located one. We observe this phenomena not only at the ``global level'' (the whole collection), but also at the ``local level'' (across the topics). \hl{In all topics, Latvian is more dominant than expected.}

@@

Over the period from 15 April 2017 to \hl{16 July 2018} the initial set of \hl{1\,959\,695} tweets was collected from the \texttt{POST status/filter} endpoint of the Twitter Streaming API. Following \cite{milajevs:2017:BUCC}, the \texttt{locations} parameter was set to the bounding box of Riga, the capital of Latvia. It is small enough to fit into Twitter API restrictions and covers about 40\% of the population of Latvia.\footnotemark{}
%
\footnotetext{The coordinates of the bounding box are \texttt{23.9325829, 56.8570671, 24.3247299, 57.0859184}.}
% IMS.2 according to http://worldpopulationreview.com/countries/latvia-population/, the population of Riga is less than half the total population of Latvia.  However, Twitter only allows bounding boxes for location to be so big, so you can't bound the whole country.

@@

To comply with Twitter's terms of service, on \hl{16 July 2018}, the raw tweet data was re-downloaded to get rid of deleted tweets. The tweets that originated a retweet were added to the collection. Also, we noticed a large number of tweets that came from Sweden (probably because of the imprecise \texttt{locations} parameter value), so the tweets with the location country code \texttt{SE} were omitted. This resulted in \hl{1\,415\,984} tweets that formed the final collection presented here.\footnotemark{}

@@

The topics are:\footnotemark{}%
\footnotetext{The topic file is available as \sfile{topics.json.txt}. Relevance judgments are available as \sfile{relevance\_judgments.csv}.}%
\begin{itemize}
\item \texttt{LV001:} \textbf{Latvia 100} \ The Centennial Celebration of the Republic of Latvia.
\item \texttt{LV002:} \textbf{Ice Hockey} \ Ice hockey.
\item \texttt{LV003:} \textbf{Refugees} \ Refugee crisis in Europe, attitude to immigration and immigrants.
\item \texttt{LV004:} \textbf{Brexit} \ Withdrawal of the United Kingdom from the European Union known as Brexit.
\item \texttt{LV005:} \textbf{Olympics} \ The Olympic games and the Latvian team.
\item \texttt{LV006:} \textbf{Winter} \ Winter, snow and cold weather.
\item \texttt{LV007:} \textbf{May 4} \ The Restoration of Independence Day.
\item \texttt{LV008:} \textbf{May 8-9} \ Remembrance of the end of Word War II (May 8). It is also honored as the Victory Day on May 9. The Europe Day is observed on May 9. Tweets about any of the three events are relevant.
\item \texttt{LV009:} \textbf{Midsummer} \ Midsummer celebration on June 23-24.
\item \texttt{LV010:} \textbf{November 11} \ A memorial day for soldiers who fought for the independence of Latvia.
\item \texttt{LV011:} \textbf{November 18} \ Proclamation Day of the Republic of Latvia.
\item \texttt{LV012:} \textbf{Christmas} \ Christmas.
\item \texttt{LV013:} \textbf{New Year} \ New Year.
\item \texttt{LV014:} \textbf{March 16} \ Remembrance Day of the Latvian legionnaires.\footnotemark{}
\item \texttt{LV015:} \textbf{The Chronicles of Melanie} \ Mel\=anijas Hronika (The Chronicles of Melanie) is a Latvian movie that was selected for the Foreign-Language Category for the Oscars.
\item \texttt{LV016:} \textbf{Blizzard of Souls} \ A Latvian movie Dv\=ese\c{l}u Putenis (Blizzard of Souls) that is in production by the time of writing.
\end{itemize}
